---
output:
  word_document: default
  html_document: default
---
#'----------------------
#'
#' Homework week 03
#'
#' Date 2018-09-28
#'
#' Laura Howes
#' 
#' ---------------------
#' 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 1

```{r eval=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
```
###10a

```{r eval=FALSE}
setwd("C:/Users/Laura/Dropbox/Grad school/BIOL 607 Biostats/Homework/data")
getwd()

Gene_data <- read.csv("./04q11NumberGenesRegulated.csv")
Gene_data
str(Gene_data)

Mean_Gene_data <- mean(Gene_data$ngenes)
SD_Gene_data <- sd(Gene_data$ngenes)
#####n = 109
SE_Gene_data <- (SD_Gene_data/sqrt(109))
CI_Gene <- (2*SE_Gene_data)
CI_Lower_Gene <- Mean_Gene_data-(CI_Gene)
CI_Lower_Gene
CI_Upper_Gene <- Mean_Gene_data+(CI_Gene)
CI_Upper_Gene
```

#####The confidence interval is 12.71 < Mean_Gene_Data < 16.29, or you could write it as 14.5 +or- 1.79

###10b

#####This confidence interval from 12.71 to 16.29 shows that in 95% of samples of the sample population will contain the true mean of the population (within that range).


###17

#####This the correct interpretation, as 0.9 < mean < 3.1 as a confidence interval states 95% of samples will contain the true population mean, and the remaining 5% of samples (2.5% each) could fall above 0.9 or 3.1.


###18a-f data set-up

```{r eval=FALSE}
beetles_per_night <- c(51,45,61,76,11,117,7,132,52,149)
```
###18a

```{r eval=FALSE}
beetles_per_night_mean <- mean(beetles_per_night)
beetles_per_night_mean
```
#####mean is 70.1

```{r eval=FALSE}
beetles_per_night_sd <- sd(beetles_per_night)
beetles_per_night_sd
```
#####SD is 48.50074

###18b
```{r eval=FALSE}
beetles_per_night_SE <- beetles_per_night_sd/sqrt(length(beetles_per_night))
beetles_per_night_SE
```
#####SE is 15.33728

###18c
```{r eval=FALSE}
CI_beetles <- 2*beetles_per_night_SE
CI_beetles
```
#####CI = 30.67456

```{r eval=FALSE}
CI_lower <- beetles_per_night_mean - (2*CI_beetles)
CI_lower
```
#####CI lower = 8.750872
```{r eval=FALSE}
CI_upper <- beetles_per_night_mean + (2*CI_beetles)
CI_upper
```
#####CI upper = 131.4491

###18d 
####If I had 25 points instead of 10, would the Mean be greater, less, or the same? 
#####The mean would be about the same no matter what sample size

###18e 
####If I had 25 points instead of 10, would the SD be greater, less, or the same? 
#####The SD would be about the same as well, as it's a measure of population like the mean

###18f 
####If I had 25 points instead of 10, would the SE be greater, less, or the same? 
#####The standard error would probably be smaller, since more data would help with precision and you would be dividing by a larger value (since you're taking the sqaure root of a bigger number)

# Problem 2

###2.1 
####Load the data using readr and make the Month_Names column into a factor whose levels are in order of month using  forcats::fct_inorder. Use levels() - a function that takes a factor vector and returns the unique levels - on the column. Are they in the right order?

```{r eval=FALSE}
library(readr)
sea_ice_data_readr <- read.csv("./NH_seaice_extent_monthly_1978_2016.csv")

sea_ice_data_readr <- sea_ice_data_readr %>% 
  mutate(Month_Name = factor(Month_Name)) %>%
  mutate(Month_Name = fct_inorder(Month_Name)) 

levels(sea_ice_data_readr$Month_Name)
```

#####They are not in the right order, just the ordered they were listed in the data "Nov" "Dec" "Feb" "Mar" "Jun" "Jul" "Sep" "Oct" "Jan" "Apr" "May" "Aug"

###2.2 
####Try fct_rev() on ice$Month_Name (I called my data frame ice when I made this). What is the order of factor levels that results? Try out  fct_relevel(), and last, fct_recode() as well. Look at the help files to learn more, and in particular try out the examples. Use these to guide how you try each functino out. After trying each of these, mutate month name to get the months in the right order, from January to December. Show that it worked with levels()

```{r eval=FALSE}
fct_rev(sea_ice_data_readr$Month_Name)
```
#####They are in reverse order from the levels function. Aug May Apr Jan Oct Sep Jul Jun Mar Feb Dec Nov

```{r eval=FALSE}
?fct_relevel
sea_ice_relevel <- fct_relevel(sea_ice_data_readr$Month_Name,"Jan","Feb","Mar","Apr","May",
                               "Jun","Jul","Aug","Sep","Oct","Nov","Dec")
sea_ice_relevel

?fct_recode
sea_ice_recode <- fct_recode(sea_ice_data_readr$Month_Name,"Jan","Feb","Mar","Apr","May","Jun",
                                         "Jul","Aug","Sep","Oct","Nov","Dec")
sea_ice_recode

sea_ice_mutated <- sea_ice_data_readr %>%
  mutate(Month_Name = fct_relevel(Month_Name,"Jan","Feb","Mar","Apr","May",
                                  "Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
sea_ice_mutated

levels(sea_ice_mutated$Month_Name)
```

###2.3 
####Now, using what you have just learned about forcats, make a column called Season that is a copy of Month_Name. Use the function  fct_recode to turn it into a factor vector with the levels Winter, Spring, Summer, Fall in that order. Use levels() on ice$Season to show that it worked.

```{r eval=FALSE}
sea_ice_mutated <- sea_ice_mutated %>%
  mutate(Season = Month_Name) %>%
  mutate(Season = fct_recode(Season, Winter = "Jan", Winter = "Feb", Spring = "Mar",
                             Spring = "Apr", Spring = "May", Summer = "Jun", 
                             Summer = "Jul", Summer = "Aug", Fall = "Sep", Fall ="Oct",
                             Fall= "Nov", Winter = "Dec"))

sea_ice_mutated

levels(sea_ice_mutated$Season)
```

###2.4a 
####Make a boxplot showing the variability in sea ice extent every month.

```{r eval=FALSE}
Sea_ice_extent <- ggplot(data = sea_ice_mutated,
                         mapping = aes(x = Month_Name, y = Extent))

Sea_ice_extent +
  geom_boxplot()
```

###2.4b 
####Use dplyr to get the annual minimum sea ice extent. Plot minimum ice by year, and add a trendline (either a smooth spline or a straight line).

```{r eval=FALSE}
Sea_ice_ann_min <- sea_ice_mutated %>%
  group_by(Year) %>%
  summarize(min_ice = min(Extent))

ggplot(data = Sea_ice_ann_min,
       mapping = aes(x = Year, y = min_ice)) +
       geom_point() +
       stat_smooth(method = "auto")
```

###2.5 
####With the original data, plot sea ice by year, with different lines for different months. Then, use facet_wrap and cut_interval(Month, n=4) to split the plot into seasons.

```{r eval=FALSE}
Sea_ice_by_year <- ggplot(data = sea_ice_mutated, 
                          mapping = aes(x = Year, y = Extent,
                                        group = Month, color = Month))

Sea_ice_by_year_plot <- Sea_ice_by_year +
  geom_line() +
  scale_color_gradientn(colors = rainbow(12)) +
  facet_wrap(~Season)

Sea_ice_by_year_plot
```

###2.6 
####Last, make a line plot of sea ice by month with different lines as different years. Gussy it up with colors by year, a different theme, and whatever other annotations, changes to axes, etc., you think best show the story of this data. For ideas, see the lab.

```{r eval=FALSE}
library(viridisLite)
library(viridis)

Line_plot_Sea_Ice_Year <- ggplot(data = sea_ice_mutated,
                                 mapping = aes(x = Month_Name, y = Extent,
                                               group = Year, color = Year))

Line_plot_Sea_Ice_Year +
  geom_line() +
  scale_color_viridis(option = "A") +
  ggtitle("Sea Ice by Year")
```


###2.7 
#####I couldn't get gganimate to install for me :(

###2.8 something new

```{r eval=FALSE}
Sea_ice_by_year_jitter <-ggplot(data = sea_ice_mutated,
                                mapping = aes(x = Year, y = Extent, 
                                              group = Month, color = Month))
  
Sea_ice_by_year_jitter +
  geom_jitter()
  scale_color_gradient(low = "blue", high = "red")
```


###EC attempt to install gganimate:

```{r eval=FALSE}
install.packages('devtools')
devtools::install_github('thomasp85/gganimate')

devtools::install_github("thomasp85/tweenr")
devtools::install_github("thomasp85/transformr")
```
